Stochastic blockmodeling of relational event dynamics

### 2012-05-31 docs/inference.tex
Add caption to intro figure.

We use Gibbs sampling to sample the latent class assignments $\mathbf{z}$ from the conditional distribution
\begin{align*}
p(z_r | \mathbf{z}_{-r},\mathcal{A},\alpha,\boldsymbol{\beta}) \propto&  p(\mathcal{A}|\mathbf{z},\boldsymbol{\beta}) p(z_r | \mathbf{z}_{-r},\alpha) \\
p(\mathcal{A}|\mathbf{z},\boldsymbol{\beta}) \propto &
\prod_{m=1}^M \lambda_{i_m,j_m}(t_m|\cdot)^{\mathbf{1}[r \in \{i_m,j_m\}]}  \prod_{(i,j) \in \mathcal{R}_r} \exp \{ -(t_m - t_{m-1}) \lambda_{ij}(t_m|\cdot)\}
\end{align*}
where $\mathcal{R}_r = \{(i,j) \in \mathcal{R}: r \in \{i,j\}\}$ is the set of dyads involving node $r$.
Under a CRP($\alpha$) prior, we have $p(z_i = k | z_{-i},\alpha) = n_k $ and $p(z_i = \mbox{new} |  z_{-i},\alpha) = \alpha$ where $n_k$ is the number of nodes assigned to cluster $k$.
In our experiments we use $\alpha=0.1$ and use Algorithm 8 from \cite{Neal2001} with 5 extra clusters drawn from the prior.\footnote{I also tried out a prior for cluster sizes such that $n_k \sim \mbox{NegBinom}(3,N/K)$, where $K$ is chosen a priori.  This aims for similarly sized groups.}

@@

>>>>

We use Gibbs sampling to sample the latent class assignments $\mathbf{z}$ from the conditional distribution
\begin{align*}
p(z_r | \mathbf{z}_{-r},\mathcal{A},\alpha,\boldsymbol{\beta}) \propto&  p(\mathcal{A}|\mathbf{z},\boldsymbol{\beta}) p(z_r | \mathbf{z}_{-r},\alpha) \\
p(\mathcal{A}|\mathbf{z},\boldsymbol{\beta}) \propto &
\prod_{m=1}^M \lambda_{i_m,j_m}(t_m|\cdot)^{\mathbf{1}[r \in \{i_m,j_m\}]}  \prod_{(i,j) \in \mathcal{R}_r} \exp \{ -(t_m - t_{m-1}) \lambda_{ij}(t_m|\cdot)\}
\end{align*}
where $\mathcal{R}_r = \{(i,j) \in \mathcal{R}: r \in \{i,j\}\}$ is the set of dyads involving node $r$.
Under a CRP($\alpha$) prior, we have $p(z_i = k | z_{-i},\alpha) = n_k $ and $p(z_i = \mbox{new} |  z_{-i},\alpha) = \alpha$ where $n_k$ is the number of nodes assigned to cluster $k$.

@@

\subsection{Hyperparameter settings}
In our experiments we use $\alpha=0.1$ and use Algorithm 8 from \cite{Neal2001} with 5 extra clusters drawn from the prior.\footnote{I also tried out a prior for cluster sizes such that $n_k \sim \mbox{NegBinom}(3,N/K)$, where $K$ is chosen a priori.  This aims for similarly sized groups.}
We set $\alpha_{\sigma}=3$ and $\beta_{sigma}=1$ so that in the presence of little data we encourage shrinkage towards the upper level parameters $\mu$.  
As with other models, we note that the predictive accuracy of the model can depend on the hyperparameters.

### 2012-05-31 docs/intro.tex
Add caption to intro figure.

% TODO: Mention dynamic stochastic equivalence

@@

\begin{figure}
\centering
\includegraphics[scale=.8]{../figs/introexample/all}
\caption{}
\label{fig:example}
\end{figure}

>>>>

@@

\begin{figure}
\centering
\includegraphics[scale=.8]{../figs/introexample/all}
\caption{Illustration of differentiation by rate of interaction versus by dynamics.  The left column shows an example where within-block communication is large, and there is a higher tendency for reciprocity within a block.  In the situation where the two groups are not differentiated by rate, as in the right column, a standard blockmodel is unable to distinguish between groups A and B.  The proposed method can, however, learn such groups by employing more flexible definitions of shared dynamics.}
\label{fig:example}
\end{figure}

### 2012-05-31 docs/model.tex
Add caption to intro figure.

We allow the blocks to share information by placing a hierarchical prior on the collection of $\boldsymbol{\beta}_{k.l}$ where $\mu_p \sim N(0,1)$ and $\sigma_p^2 \sim \mbox{Inv-Gamma}(3,1)$.
The cluster assignments are given a non-parameteric priors $z_i \sim \mbox{CRP}(\alpha)$.

>>>>

We allow the blocks to share information by placing a hierarchical prior on the collection of $\boldsymbol{\beta}_{k.l}$ where $\mu_p \propto 1$ and $\sigma_p^2 \sim \mbox{Inv-Gamma}(\alpha_{\sigma},\beta{\sigma})$.
The cluster assignments are given a non-parameteric priors $z_i \sim \mbox{CRP}(\alpha)$.
